\subsection{Mutations and its effects in the central dogma of molecular biology}
Within the human genome mutations occur continuously by internal and external factors that substitute, remove, insert or alter the reading frame in a nucleotide sequence. Mutations are not without consequences and can be: beneficial, benign or in most cases pathogenic because they replace a nucleotide which serves a purpose at the specific position in a sequence. Alterations in sequences might lead to a difference in ribonucleic acid (RNA) transcription rates or differences in the RNA transcript that is formed from the deoxyribonucleic acid (DNA) which both can influence the cellular machinery. Mutations outside a gene could less transcription of a protein and inside a gene it could lead to proteins that are: unstable during or after formation, perform less optimal or are not functional \cite{}.
%TRAPS, NIH
%Sickle cell disease, NIH
%Cystic fibroses, NIH

\subsection{A general concept of structural levels within proteins}
The formation of protein structures is classified in different levels, distinctions are made based on bindings and structures that arise from them. 
The order in which amino acids appear in a sequence is called the primary structure, in this level amino acids are only bound to each other by peptide bonds. 
Within a primary structure amino acids can form new peptide bonds between the N and C -terminus of an amino acid, with these bonds 3D structures are made called $\alpha$ helices and $\beta$-sheets that together make up the secondary structure.
More alterations to a single amino acid sequence in the 3D can come from disulfide bridges, ion, hydrogen -bonds, hydrophobic and hydrophilic -interactions formed by the residues of the amino acids, together these bonds form the tertiary structure.
By combining multiple tertiary structures the quaternary structure of a protein can be formed out of the mentioned bonds, bridges and interactions \cite{}.
% Protein structure, Wikipedia
%Protein Conformation and Diagnostic Tests: The Prion Protein, Bennion


\subsection{Kaas}

Acquiring information about DNA sequences highly relies on experimental sequencing methods and became cheaper over the years \cite{} and found its use in diagnosing patients within the healthcare sector \cite{}.
%The Cost of Sequencing a Human Genome, NIH
%GAVIN: Gene-Aware Variant INterpretation for medical sequencing, van der Velde 
From the collected data by genome sequencing experiments most of the analysis is handled in-silico due to the quantities of data that is produced. Proteins often find their use in diagnosing diseases experimentally \cite{}, however in-silico it is often limited to information about conservation in the amino acid sequence which may lead to identical results as by analyzing DNA.
% Introduction: Advances in Protein Analysis for the Clinical Laboratory, Hortin
%Protein Conformation and Diagnostic Tests: The Prion Protein, Bennion
Yet the 3D structure defines how a protein functions \cite{} and by assessing structures of protein variants it becomes possible to determine the change in function and diagnose protein variants that were unclassifiable through finding conservation.
%How do crystal lattice contacts reveal protein crystallization mechanism?, Nanev
Another advantage of the structural information is that it gives the possibility to develop treatment for diseases that are caused by a mutations. 
With experimental methods such as X-ray crystallography and nuclear magnetic resonance (NMR) more than 158000 structures \cite{} have been completely revealed, however it is only a tiny fraction of the potential proteins possible without folds \cite{}.
% wwPDB: Deposition Statistics, wwPDB	
%Chemiotics: How many proteins can we make? : The Sceptical Chymist, Cantrill
Making 3D structures is currently not common for diagnosis because it is relative expensive and it is difficult to perform with some structures that contain flexible regions where in the positions of atoms is hard to determine the exact position.
% %PDB101: Learn: Guide to Understanding PDB Data: Missing Coordinates and Biological Assemblies, PDB101
A cheaper alternative is modeling of proteins which starts with the primary sequence 



 are relative flexible which makes it un

which makes it difficult to determine the exact position of atoms within the structure.

Generating structures 
Difficulties of acquiring
Several problems for acquiring structures are price and even more the complexity of acquiring 

While more than 158000 \cite{} structures are complete it is still a tiny fraction of the wide span of potential proteins \cite{} that can be formed even without structural information added. 

Which finds its origin in the price of acquiring structures and the complexity, some parts of a structure are highly flexible which makes it difficult to determine the exact position within a crystal lattice structure \cite{}.




 not every protein structure has the same biochemical composition and can form different structures of which some are difficult to di

and can cause different structures to be formed 

 is identical and all have different biochemical properties resulting in structural 

which form in some protein structures that are continuously moving which makes it hard to identify the exact composition. 

One of the problems with these methods is that not every protein has parts that are motionless an
 due to the complexity and the price of acquiring the structures of proteins with methods such as X-ray crystallography and nuclear magnetic resonance (NMR).

With the quantities of generated sequences and the limits of what currently is found the addition of protein data 

With the vast amounts of methods of data generated for diagnosis 

While amino acid sequences can be translated from DNA, structures cannot and can currently only be determined by expensive experimental methods such as X-ray crystallography and nuclear magnetic resonance (NMR).
While protein information statets more about what product is formed and how it can cause a disease.
However protein information is not and determining structures is still difficult and expensive with methods such as X-ray crystallography and nuclear magnetic resonance (NMR).

%Discovering mutations in the genome is done with sequencing and is a routine practice in the healthcare sector \cite{}. With this method vast amounts of data are generated that is over flooding the capabilities in analyzing, therefore machine learning approaches have been developed to diagnosis whether a mutation is benign or pathogenic \cite{}.

Using DNA and RNA in diagnosis is becoming more common but using protein information is not, determining structures with X-ray crystallography and nuclear magnetic resonance (NMR) is expensive and does not guarantee
Although the problem of a disease might lie in the protein in se

But acquiring protein structures for healthcare is not common and is relative expensive to use for a diagnosing a problem that often can be found with different methods.  with methods such as X-ray crystallography and nuclear magnetic resonance (NMR).

 error in difficult identifiable regions such as loops.


%and is currently routine in the healthcare sector \cite{}. 
%
%\subsection{Problems in protein modeling and strategies for resolving structures}
%More than 158000 \cite{} proteins structures have been resolved with methods such as X-ray crystallography and nuclear magnetic resonance (NMR) \cite{}.
%% wwPDB: Deposition Statistics, wwPDB	
%%PDB101: Learn: Guide to Understanding PDB Data: Methods for Determining Structure, PDB101
%
%
%
%
%More than 158000 \cite{} proteins structures have been resolved with methods such as X-ray crystallography and nuclear magnetic resonance (NMR) \cite{}, which is a relative small amount compared to all publicly available DNA and RNA sequences \cite{}. 

%%PDB101: Learn: Guide to Understanding PDB Data: Methods for Determining Structure, PDB101
%GenBank and WGS Statistics, NCBI
%But 
%
%
% Both are resource intensive and are not able to determine every structure yet because  
%At the moment of writing more than 158000 \cite{} protein structures have been resolved by experimental methods such a 
%
%
%
%
%Importance of experimental determination instead of just modeling
%
%Modeling protein techniques as homology modeling and protein threading
%
%Simulations such as gromacs 
%
%
%
%
%
%Mutations can be discovered by genome sequencing which is a routine practice in healthcare sector \cite{} which leads to vast amounts of DNA sequence data that describe how genes are structured and can be analy 
%%  GAVIN: Gene-Aware Variant INterpretation for medical sequencing, van der Velde 
%
%
%In some situations it is known that a variant has a pathogenic or benign effect and
%

%
%with exact order of the DNA sequence and over flooding the human capabilities of analyzing and perceiving   with known and unknown effects is important for diagnosis of patients for determining the dise
%
%Mutations can be discovered by sequencing the genome, which is routine practice in the healthcare sector \cite{}, resulting in vast amounts of data which all need to be analyzed and can be compared to previously discovered information.  
%
%To determine the meaning of these variants machine learning software has been developed to aid in genome diagnostics to determine the effects of mutations \cite{}.
%  GAVIN: Gene-Aware Variant INterpretation for medical sequencing, van der Velde 

%Within DNA sequences of humans continuously mutations occur from internal and external factors, some of these mutation insert, remove or substitute nucleotides which alter the the rate of tr
% Some types of mutations can be repaired \cite{}, other mutations are made on purpose\cite{}, incorrectly repaired \cite{} or are irreparable by the available mechanisms \cite{} and will be carried over to the next generation of cells or even to the offspring.
%DNA repair, Wikipedia
%Janeway’s Immunobiology page 410-411 10.7: Germinal center B cells undergo V-region somatic hypermutation, and cells with mutations that improve af nity for antigen are selected, Kenneth murphy
% Repair of naturally occurring mismatches can induce mutations in flanking DNA, Chen et al.
% Error-Prone Repair of DNA Double-Strand Breaks, Rodgers and McVey


%
%Evidence is based on conservation whether it works or not
%
%Telling about the current state of gene variant prediction in the perspective of machine learning, GAVIN, SIFT, CADD

